##########################################
########### Survival Analysis ############
##########################################


################################################################################################
############ This code contains two parts: piece-wise analysis and global analysis #############
################################################################################################

```{r}
library(dplyr)
library(survival)
library(tidyr)
library(ggplot2)
library(survex)
library(randomForestSRC)
library(lubridate)
library(GGally)
library(patchwork)
library(scales)  # for gradient color scaling
```
######################## load simulated data for survival analysis
```{r}
simulated_IL2 <- read.csv("Simulated_IL2.csv")
simulated_IFNg <- read.csv("Simulated_IFNg.csv")
simulated_igG <- read.csv("simulated_igG.csv")
simulated_MBC <- read.csv("simulated_MBC.csv")
simulated_IC50 <- read.csv("simulated_IC50.csv")
```
############# Merge simulated data ###############
```{r}

merged_data <- simulated_igG %>%
  left_join(simulated_IL2, by = c("Code","Days"))%>%
  left_join(simulated_IFNg, by = c("Code","Days"))%>%
  left_join(simulated_MBC, by = c("Code","Days")) %>%
  left_join(simulated_IC50, by = c("Code","Days")) %>%
  rename(ID = Code) %>%
  rename(nAb = IC50) %>%
  mutate(
    log_IgG = log10(Anti_S_IgG),
    log_MBC = log10(Spike_MBC),
    log_nAb = log10(nAb),
    log_IFNg = log10(IFNg),
    log_IL2 = log10(IL2)
  ) 
merged_data
```
############### Read and Merge raw data #############
```{r}
raw_IL2 <- read.csv("")
raw_IFNg <- read.csv("")
raw_igG <- read.csv("")
raw_MBC <- read.csv("")
raw_IC50 <- read.csv("")

raw_merged_data <- raw_igG[,c(1:3)] %>%
  left_join(raw_IL2[,c(1:3)], by = c("ID","Day"))%>%
  left_join(raw_IFNg[,c(1:3)], by = c("ID","Day"))%>%
  left_join(raw_MBC[,c(1:3)], by = c("ID","Day")) %>%
  left_join(raw_IC50[,c(1:3)], by = c("ID","Day")) %>%
  rename(nAb = IC50) %>%
  mutate(
    log_IgG = log10(Anti.S.IgG),
    log_MBC = log10(SPIKE_MBC),
    log_nAb = log10(nAb),
    log_IFNg = log10(IFNg),
    log_IL2 = log10(IL2))
    #raw_merged_data <- na.omit(raw_merged_data)
    raw_merged_data
```



################################################################################
####################### Plot correlations, S Figure 7 ##########################
################################################################################

```{r}
cor_data <- merged_data %>%
  group_by(ID) %>% slice(218:392) %>%  ungroup() %>% dplyr::select( log_IgG, log_nAb, log_MBC, log_IFNg, log_IL2) # This example of slice:218-392 is representing 6M-12M, change it to 30-123/124-217 to obtain results for 30D-3M and 3M-6M
#cor_data <- raw_merged_data %>% dplyr::select( log_IgG, log_nAb, log_MBC, log_IFNg, log_IL2) # This line is for S-Fig-7-1

# Custom correlation panel with color
cor_fun_colored <- function(data, mapping, method = "spearman", ...){
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  
  corr <- cor(x, y, method = method, use = "complete.obs")
  p_val <- cor.test(x, y, method = method)$p.value
  
  stars <- symnum(p_val, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 1),
                  symbols = c("***", "**", "*", ""))
  
  # Use absolute correlation to define color intensity
  col <- gradient_n_pal(c("#FFE5CC", "#FF6600", "#990000"))(abs(corr))  # light to strong red/orange
  
  ggally_text(
    label = paste0("Corr:\n", sprintf("%.3f", corr), stars),
    mapping = aes(),
    color = col,
    size = 5,
    ...
  )
}
cor_plot <- ggpairs(
  cor_data,
  lower = list(continuous = wrap("points", alpha = 0.5, size = 0.4)),
  upper = list(continuous = cor_fun_colored),
  diag = list(continuous = wrap("densityDiag", color = "blue"))
) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    panel.spacing = unit(0, "lines"),
    strip.background = element_rect(fill = "gray90", color = NA),
    strip.text = element_text(size = 10, face = "bold", color = "black"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 10)
  )

print(cor_plot)
```

#######################################################
########## Immunity details of individual #############
#######################################################
```{r}
immu_detail_individuals <- read.csv("immu_detail_individuals_reinf.csv")
```
```{r}
# retrieve individual infection info 
Infection_info <- immu_detail_individuals %>%
  dplyr::select(ID,inf,t_inf, reinf, t_inf_2) #%>%
Infection_info
```
```{r}
# merge simulated data with inf_info
merged_data_immu <- merged_data %>%
  left_join(Infection_info, by = "ID"  )
merged_data_immu
```
```{r}
# merge raw data with immunity info and compute max day of visit for each
full_data_withoutVIV_immu <- merge(raw_merged_data, immu_detail_individuals, by = c("ID")) %>% group_by (ID) %>% mutate(Visit = row_number() )
ind_max_visit <- full_data_withoutVIV_immu %>% group_by(ID) %>% slice_tail() %>% dplyr::select(ID,Day,Visit)
ind_max_visit
```
```{r}
# Compute the mean day of each visit. Based on this, we knew 3M corresponds to 123D, 6M~217D. 12M~392D
mean_day <- full_data_withoutVIV_immu %>% group_by(Visit) %>%
  summarize(Avg_Day = median(Day, na.rm = TRUE),
            Q25 = quantile(Day, 0.25, na.rm = TRUE),
    Q75 = quantile(Day, 0.75, na.rm = TRUE))
mean_day
```

##################################################################################################################################
################################## Piecewise Analysis: 30D-3M, 3M-6M, 6M-12M #####################################################
##################################################################################################################################

# Process time scale 
```{r}
# For those uninfected, the time cuts off at the maximum visit date
uninf_ind_max_followup <- full_data_withoutVIV_immu %>%
  filter(inf == 0) %>%
  group_by(ID) %>%
  slice_tail(n=1) %>%
  dplyr::select(ID, Day)
uninf_ind_max_followup
```
```{r}
# For infected, cutoff = time of infection
# For uninfected, cutoff = time of last visit
# For reinfected, cutoff = time of reinfection

inf_ind_time <- full_data_withoutVIV_immu %>%
  filter(inf == 1& reinf ==0) %>%
  group_by(ID) %>%
  mutate(time = t_inf) %>%
  slice_head(n=1) %>%
  dplyr::select(ID, time) %>%
  rename(Day = time) #infected ones has t_max = t inf
inf_ind_time

boost_ind_time <- uninf_ind_max_followup %>%
  filter(ID %in% subset(full_data_withoutVIV_immu,boost == 1&inf == 0)$ID)
boost_ind_time

uninf_ind_max_followup<- uninf_ind_max_followup %>%
  filter(!ID %in% boost_ind_time$ID)
uninf_ind_max_followup  

reinf_ind_time <- full_data_withoutVIV_immu %>%
  filter(reinf == 1) %>%
  group_by(ID) %>%
  mutate(time1 = t_inf,
         time2 = t_inf_2) %>%
  slice_head(n=1) %>%
  dplyr::select(ID, time1,time2) %>%
  rename(Day = time1) #infected ones has t_max = t inf
reinf_ind_time
```
```{r}
#Combine all scenarios 
max_followup <- rbind(uninf_ind_max_followup,inf_ind_time,reinf_ind_time) %>%
  rbind(boost_ind_time)%>%
  rename(max_t = Day)
max_followup
```
```{r}
# Combine the infection information with maximum followup
inf_info <- immu_detail_individuals[,c(1,2,6)] %>%
  left_join(max_followup, by = "ID" ) 
inf_info  

######## For df inf_info, "max_t" is the first time infection date or max visit date, "time2" is the reinfection date 
```

####### Start performing  survival analysis 
```{r}
###############
head = 30  ### define the starting date of piecewise analysis: 30, 124, 218
tail = 123 ### define the ending date of piecewise analysis: 123, 217, 392

inf_b4_n <- inf_info %>% 
  filter(inf==1 & max_t <head)  ### find out who gets infection before the starting date: as those are defined as hybrid immunity

merged_data_immu_30 <- merged_data_immu %>%
  group_by(ID) %>%
  slice(head:tail) %>%
  ungroup() ### Slice the simulated data within a certain period

###############

# redefine the infection events based on the study period
new_inf_sub1 <- inf_info %>% filter(reinf ==0)%>%
  mutate(new_inf = ifelse(inf==1 &max_t>=head&max_t<=tail,1,0)) %>%
  mutate(new_max_t = ifelse(new_inf==1,max_t,tail)) 
new_inf_sub2 <- inf_info %>% filter(reinf ==1)%>%
  mutate(new_inf = ifelse((max_t>=head&max_t<=tail)|(time2>=head&time2<=tail),1,0)) %>%
  mutate(new_max_t = ifelse(new_inf==1,max_t,tail)) %>%
  mutate(new_max_t = ifelse(time2>=head&time2<=tail, time2, new_max_t)) %>%
  mutate(new_max_t = ifelse(head< max_t&tail>max_t, max_t, new_max_t)) 

new_inf <- rbind(new_inf_sub2,new_inf_sub1)%>%
  dplyr::select(ID,new_max_t,inf,new_inf)


count<- new_inf %>%
  #filter(! ID %in% inf_b4_n$ID) %>%
  group_by(new_inf) %>%
  summarise(count = n()) 
count_original <- new_inf %>%
  group_by(inf) %>%
  summarise(count = n()) 

####### Create the survival dataset for the time-varying covariate Cox-Ph model
df<-tmerge(new_inf,new_inf,id=ID,
endpt=event(new_max_t,new_inf))

df_processed <- tmerge(df,merged_data_immu_30,id=ID, # First argument: df: primary dataset, Second argument: dataset contians the covariates
 log_Anti_S_IgG=tdc(Days,log_IgG),
 log_Spike_MBC = tdc(Days, log_MBC),
 log_nAb = tdc(Days, log_nAb),
  log_IFNg = tdc(Days, log_IFNg),
 log_IL2 = tdc(Days, log_IL2),
 Anti_S_IgG=tdc(Days,Anti_S_IgG),
 IL2 = tdc(Days, IL2),
 Spike_MBC = tdc(Days, Spike_MBC),
 IC50 = tdc(Days, IC50),
 IFNg = tdc(Days, IFNg)) %>% na.omit() %>%
  mutate(
    scale_IgG = scale(log_Anti_S_IgG),
    scale_MBC = scale(log_Spike_MBC),
    scale_nAb = scale(log_nAb),
    scale_IFNg = scale(log_IFNg),
    scale_IL2 = scale(log_IL2)
 )
df_processed <- na.omit(df_processed)

# df_processed <- df_processed %>% filter(!ID %in% inf_b4_n$ID ) ############# Run this line if you target children withou hybrid immunity ##############
df_processed

############ Start fitting
fit.tdc <- coxph(Surv(tstart,tstop,endpt)~   # Choose the biomarker you want to fit
   # log_Anti_S_IgG +
   # log_nAb+
   # log_Spike_MBC+
   # log_IFNg+
   # log_IL2+
   #   scale_IgG+
   #   scale_nAb +
   #   scale_MBC +
   #   scale_IFNg +
   #   scale_IL2 +

    cluster(ID)
  ,df_processed,x = TRUE)

summary(fit.tdc)
```



##################################################################################################################################
################################## Global Analysis: Children without hybrid immunity #############################################
##################################################################################################################################


## for merged simulated data, filter children who got infection after day 30
```{r}
b4_hybrid <- subset(inf_info,max_t>=30) %>% dplyr::select(ID,inf, max_t) 
b4_hybrid 
count<- b4_hybrid %>%
  group_by(inf) %>%
  summarise(count = n()) 
count # n =69
```
################### Based on the starting and ending of each child, slice the data
```{r}
b4_hybrid_data <- list()
for (i in seq_along(b4_hybrid$ID)) {
  id <- b4_hybrid$ID[i]
  cut_off <- b4_hybrid$max_t[i]
  ind_data <- merged_data %>% filter(ID == id) %>% slice(30:cut_off)
  b4_hybrid_data[[i]] <- ind_data
}

b4_hybrid_data <- bind_rows(b4_hybrid_data) 
```
################### Create the survival dataset for the time-varying covariate Cox-Ph model
```{r}
# process the event information
df<-tmerge(b4_hybrid,b4_hybrid,id=ID,
endpt=event(max_t,inf))

df_processed <- tmerge(df,b4_hybrid_data,id=ID, # First argument: df: primary dataset, Second argument: dataset contians the covariates
 log_Anti_S_IgG=tdc(Days,log_IgG),
 log_Spike_MBC = tdc(Days, log_MBC),
 log_nAb = tdc(Days, log_nAb),
  log_IFNg = tdc(Days, log_IFNg),
 log_IL2 = tdc(Days, log_IL2),
 Anti_S_IgG=tdc(Days,Anti_S_IgG),
 IL2 = tdc(Days, IL2),
 Spike_MBC = tdc(Days, Spike_MBC),
 nAb = tdc(Days, nAb),
 IFNg = tdc(Days, IFNg)) %>% na.omit() %>%
  mutate(
    scaled_IgG = scale(log_Anti_S_IgG),
    scaled_MBC = scale(log_Spike_MBC),
    scaled_nAb = scale(log_nAb),
    scaled_IFNg = scale(log_IFNg),
    scaled_IL2 = scale(log_IL2)
 ) 
df_processed <- na.omit(df_processed)
df_processed
```
########### Start fitting ############
```{r}
fit.tdc <- coxph(Surv(tstart,tstop,endpt)~
 #   log_Anti_S_IgG + 
 #   log_nAb+
 #   log_Spike_MBC+
 #   log_IFNg+
 #   log_IL2+ 
  #  scaled_IgG+
    scaled_nAb +
    scaled_MBC +
    scaled_IFNg +
  #  scale_IL2 +
   cluster(ID),data = df_processed)
summary(fit.tdc)
```

########## Test the proportionality ###########

```{r}
zph <- cox.zph(fit.tdc)
zph
par(mfrow = c(1, 3))   # 3 panels side by side
plot(zph, lwd = 2, col = 1:3)   # cox.zph object with 3 covariates
```

########## Plot S figure 9 #############
```{r}
beta_full <- list()
zph <- cox.zph(fit.tdc)
for (i in 1:n_distinct(colnames(zph$y))) {
  
  
  
  zph_identity <- cox.zph(fit.tdc   , transform = "identity")
  plot_time <- plot(zph_identity[i], plot = FALSE)
  
  plot_i <- plot(zph[i], plot = FALSE)
  beta_df=plot_i$y
   
  df <- data.frame(
    time = plot_time$x,
    beta = beta_df[,1],
    lower_ci=beta_df[,3],
    upper_ci=beta_df[,2],
    covariate = colnames(zph$y)[i]   # store covariate name
  )
  
  beta_full[[i]] <- df
}

# Combine all into one data frame for ggplot
beta_full_df <- do.call(rbind, beta_full)
beta_full_df
```
```{r}
plot_beta <- ggplot(beta_full_df, aes(x = time, group = covariate)) +

  geom_hline(yintercept = 0, linetype = "dotted", color = "green3", linewidth = 1) +
  geom_ribbon(aes(ymin=lower_ci, ymax = upper_ci, fill = covariate), alpha = 0.1) +
  geom_line(aes(y = beta, color = covariate), size = 1) +
  theme_minimal() +
  labs(x = "Days after primary vaccination", y = " Time-varying coefficients β(t)") +
  scale_color_manual(values = c(
    "scaled_IFNg" = "#1f77b4",
    "scaled_MBC" = "#2ca02c",
    "scaled_nAb" = "#d62728",
    "scaled_IgG" = "#d62728"
  ))+
  scale_fill_manual(values = c(
  "scaled_IFNg" = "#1f77b4",
  "scaled_MBC" = "#2ca02c",
  "scaled_nAb" = "#d62728",
  "scaled_IgG" = "#d62728"
)) +
  theme(
    axis.text = element_text(colour = "black", size = 12),
    axis.ticks = element_line(colour = "black"),
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    strip.text = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "cm")
  )# +
 # scale_x_continuous(breaks = seq(30, 300, by = 50))
  
plot_beta
```

########### Since nAb violates proportionality, we want to test whether the time-dependent association of nAb is significant or not ###########
```{r}
fit.tdc <- coxph(Surv(tstart,tstop,endpt)~

   tt(log_nAb) +
   tt(log_Spike_MBC) +
   tt(log_IFNg) +
   
  #  scale_IgG+
  #  scale_MBC +
  #  scale_nAb +
  #  scale_IFNg +
  #  scale_IL2 +
    
  # cluster(ID),data = df_processed,tt = function(x, t, ...) x * (t >= 100 ))  # Define the tt function
    cluster(ID),data = df_processed,tt = function(x, t, ...) x * (t >= 30 & t <= 120))
summary(fit.tdc)
```
####### END #######
